An Algorithm For Spatial Mapping Using a Hexagon Tile Map, With Application to Australian Maps

### 2020-01-14 kobakian_sugarbag_jss/kobakian_sugarbag_jss.Rmd
future work

The Australian Cancer Atlas [@atlas] allows estimates derived from the models of Standardised Incidence Rates and excess deaths to be downloaded. 
Figure \ref{fig:melanoma-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Rate of melanoma cancer for all persons for each SA2. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Brisbane stand out as more orange and red.
Comparatively, the hexagon tile map display in Figure \ref{fig:melanoma-hex} draws attention to contrast of the blue areas in Sydney and Melbourne and the capital city of Brisbane. In both Sydney and Melbourne, the hexagons that represent the SA2 areas in the inner-city areas have lower than average Incidence Rates.

@@

```{r melanoma-geo, fig.cap = "A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated Standardised Incidence Rate of Melanoma for all persons, they range from much lower than average (blue) to much higher than average (red)", results = "asis", echo = FALSE, out.width = "60%"                             }
knitr::include_graphics("figures/aus_melanoma_p.png")
```

@@

The \pkg{gganimate} package can be used to make an animation.
It requires connecting the two displays, which can be done using the \code{sf_id} such as the SA2 name.
The animation^[This animation can be viewed at: ] connecting these two displays will highlight the rapid growth of the inner-city areas, and will decrease the large rural areas.

@@

Structure of conclusions: (1) summarise paper, discuss limitations, suggest future directions

@@

These hexagons equally represent each area. However, the tesselation does not allow the size of the hexagons to represent another variable.

>>>>

The Australian Cancer Atlas [@TACA] allows estimates derived from the models of Standardised Incidence Rates and excess deaths to be downloaded. 
Figure \ref{fig:melanoma-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Rate of melanoma cancer for all persons for each SA2. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Brisbane stand out as more orange and red.
Comparatively, the hexagon tile map display in Figure \ref{fig:melanoma-hex} draws attention to contrast of the blue areas in Sydney and Melbourne and the capital city of Brisbane. In both Sydney and Melbourne, the hexagons that represent the SA2 areas in the inner-city areas have lower than average Incidence Rates.

@@

```{r melanoma-geo, fig.cap = "A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated Standardised Incidence Rate of Melanoma for all persons, they range from much lower than average (blue) to much higher than average (red)", results = "asis", echo = FALSE, out.width = "60%"}
knitr::include_graphics("figures/aus_melanoma_p.png")
```

@@

The \pkg{gganimate} package can be used to make an animation.
It requires connecting the polygons for each area in two displays two displays, which can be done using the \code{sf_id} variable, such as the SA2 name.
The animation^[This animation can be viewed at: https://sugarbagjss.netlify.com/] connecting these two displays will highlight the rapid growth of the inner-city areas, and will decrease the large rural areas. The hexagons that move the furthest will move rapidly in the animation.

@@

@@

These hexagons equally represent each area. However, the tesselation does not allow the size of the hexagons to represent another variable, similar to the choropleth maps.
The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. It works on the assumption that viewers can use directional assumptions to identify their neighbourhoods, this can be aided by the animation.

<!-- future -->
Future work will include refining the algorithm. It would be possible to take a logarithmic function rather than a direct angle to help choose a closer hexagon to the original centroid location, before increasing the width of the angle used to filter the hexagons.

This algorithm has only been tested using single countries, and does not consider definite borders of countries. While the buffer allows extention beyond the furthest centroids, there is no mechanism to protect the borders and ensure centroids are placed within the geographic borders of a country.

This algorithm is an effective start to creating hexagon tile maps for many geographic units.

### 2020-01-14 kobakian_sugarbag_jss/kobakian_sugarbag_jss.tex
future work

XXX Show whole Australia male liver cancer with choropleth and hexagon
tile map, and discuss the pros/cons of the result, for understanding the
incidence.

\begin{CodeInput}
R> knitr::include_graphics("figures/aus_liver_m.png")
\end{CodeInput}

\begin{center}\includegraphics[width=41.67in]{figures/aus_liver_m} \end{center}

\begin{CodeInput}
R> knitr::include_graphics("figures/aus_liver_m_hex.png")
\end{CodeInput}

\begin{center}\includegraphics[width=41.67in]{figures/aus_liver_m_hex} \end{center}

\hypertarget{election-results}{%
\subsection{Election results}\label{election-results}}

XXX Code from web site, and a couple of paragraphs of explanation

@@

XXX show how to make an animation, and make a link to a web page where
the animation can be viewed

@@

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \citet{sugarbag} package written for \citet{R}. The \pkg{sugarbag}
package creates tesselated hexagon tilegrams. The Australian application
preserves the spatial relationships, emphasising capital cities. The
hexagon tile map is a visualisation solution that highlights spatial
distributions.

@@

The code for sugarbag \citep{sugarbag} can be found on
\href{https://cran.r-project.org/web/packages/sugarbag/index.html}{CRAN},
along with vignettes on installation and usage.

>>>>

The Australian Cancer Atlas \citep{TACA} allows estimates derived from
the models of Standardised Incidence Rates and excess deaths to be
downloaded. Figure \ref{fig:melanoma-geo} is a choropleth map that uses
colour to display the estimated Standardised Incidence Rate of melanoma
cancer for all persons for each SA2. The Australian choropleth map
display draws attention to the expanse of light blue areas across the
rural communities in all states. The SA2s around Brisbane stand out as
more orange and red. Comparatively, the hexagon tile map display in
Figure \ref{fig:melanoma-hex} draws attention to contrast of the blue
areas in Sydney and Melbourne and the capital city of Brisbane. In both
Sydney and Melbourne, the hexagons that represent the SA2 areas in the
inner-city areas have lower than average Incidence Rates.

With careful consideration of the choropleth map, the small geographic
inner city areas may have been noticed by viewers, but the hexagon tile
map display emphasises them. The communities in northern Queensland and
the Northern territory do not draw attention because of their size as in
the choropleth, but their colour is still noticably below average when
contrasted with the heaxgons further south.

To create this choropleth map the SA2 polygons for 2011 from the
Australian Bureau of Statistics. The Standardised Incidence Ratios for
each geographic unit are joined to the appropriate polygons.

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{figures/aus_melanoma_p}

\caption[A choropleth map of the Statistical Areas of Australia at Level 2]{A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated Standardised Incidence Rate of Melanoma for all persons, they range from much lower than average (blue) to much higher than average (red)}\label{fig:melanoma-geo}
\end{figure}

To create the hexagon tile map display, the same steps are followed as
outlined above:

\begin{itemize}
\tightlist
\item
  Create the set of centroid points
\item
  Create the hexagon grid points
\item
  Allocate each centroid to a hexagon grid point
\end{itemize}

\begin{figure}

{\centering \includegraphics[width=0.6\linewidth]{figures/aus_melanoma_p_hex}

\caption[A hexagon tile map of the Statistical Areas of Australia at Level 2]{A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated Standardised Incidence Rate, they range from much lower than average (blue) to much higher than average (red)}\label{fig:melanoma-hex}
\end{figure}

@@

The \pkg{gganimate} package can be used to make an animation. It
requires connecting the polygons for each area in two displays two
displays, which can be done using the \code{sf_id} variable, such as the
SA2 name. The animation\footnote{This animation can be viewed at:
  \url{https://sugarbagjss.netlify.com/}} connecting these two displays
will highlight the rapid growth of the inner-city areas, and will
decrease the large rural areas. The hexagons that move the furthest will
move rapidly in the animation.

@@

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \pkg{sugarbag} \citet{sugarbag} package written for \code{R}
\citet{R}. The \pkg{sugarbag} package creates tesselated hexagon
tilegrams. The Australian application preserves the spatial
relationships, emphasising capital cities. The hexagon tile map is a
visualisation solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tesselation
does not allow the size of the hexagons to represent another variable,
similar to the choropleth maps. The algorithm is heavily dependent on
the focal points used, as this determines the order of allocation. It
works on the assumption that viewers can use directional assumptions to
identify their neighbourhoods, this can be aided by the animation.

Future work will include refining the algorithm. It would be possible to
take a logarithmic function rather than a direct angle to help choose a
closer hexagon to the original centroid location, before increasing the
width of the angle used to filter the hexagons.

This algorithm has only been tested using single countries, and does not
consider definite borders of countries. While the buffer allows
extention beyond the furthest centroids, there is no mechanism to
protect the borders and ensure centroids are placed within the
geographic borders of a country.

This algorithm is an effective start to creating hexagon tile maps for
many geographic units.

@@

The code for \pkg{sugarbag} \citep{sugarbag} can be found on
\href{https://cran.r-project.org/web/packages/sugarbag/index.html}{CRAN},
along with vignettes on installation and usage.

Australian Electoral Data can be found in the \pkg{eechidna} package for
R.
